Manuscript title here

### 2021-08-08 docs/ms/ms.RMD
update befre discussion with CRM

source(here::here("datascript.R"))
source(here::here("analyses/trawl/R/Lorenz.R"))

@@

Anthropogenic activities continue to exert long-term, varied, and increasing pressures on ecosystems causing alterations to biological diversity on broad and local scales. Shifts in biodiversity are variable, nuanced and difficult to predict. Yet, understanding how communities change through time is crucial for predicting current and future changes to biological systems and the services they provide. Explaining the extent and pace of biodiversity change will require identifying the conditions that lead to divergent biological responses, e.g., whether change happens gradually or abruptly and through individual versus cohesive species responses. In particular, disentangling the roles of extrinsic, environmental forcing and intrinsic, ecological processes is vital. We propose to combine two time series of environmental and biological change with information on species’ traits and trophic interactions to attribute biological change to its extrinsic and intrinsic drivers. The proposed work includes analysis and continuation of a data set extending nearly two decades of pelagic megafaunal communities and environment monitoring
for a shallow estuarine bay along the Gulf of Mexico Coast. To this, we add analysis and continuation of a three year dataset of benthic macrofaunal from the same bay. The project proposes to address four key questions.

@@

<!--- warming and effects in food webs --->
Anthropogenic activities continue to exert long-term, varied, and increasing pressures on ecosystems [@steffen2015] causing alterations to biological diversity on local and global scales. Shifts in biodiversity at local scales are variable and potentially more nuanced than at global scales, but generally local changes manifest primarily as changes to community composition and structure rather than declines in richness [@hillebrand2018]. These shifts in communities through time are difficult to predict, in part, due to the multidimensional nature of biodiversity [@chase2018]. Yet, understanding how communities change through time is crucial for predicting current and future changes to biological systems and ecosystem services.
<!--- Speeding up of food web: Species levels--->
Explaining the extent and pace of biodiversity change will require identifying the conditions that lead to divergent biological responses to environmental change, both natural and anthropogenic. In particular, it will be important to disentangle the roles of intrinsic (e.g., ecological) and extrinsic (e.g., environmental) drivers of biodiversity change. Intrinsic and extrinsic processes can interact, leading to temporal patterns of biodiversity change that vary in response to environmental change. For example, gradual environmental change over time can lead to gradual community shifts or these shifts may be undetectable or abrupt [@williams2011][@scheffer2003].

@@

Here, we quantified over a decade of change in a shallow estuarine community. We

@@

Trawl sampling was conducted as part of the Louisiana Universities Marine Consortium's (LUMCON) educational outreach program (https://www.lumcon.edu/education-and-outreach). Individual trawls were conducted using *trawl apparatus description...* for 10 - 20 min. When multiple trawls were conducted on a single day, species counts were summed across all trawls and

@@

Species names were checked for typos and confirmed using the *gni_parse* function from the *taxize* package [@chamberlain2020] and further standardized to consistent taxonomic resolution for entire data series (Supplemental materials). We measured multiple aspects of species richness, first to track the uncorrected accumulation of species over time we filtered species from the data set after first occurrence. We then aggreagated all trawls within a calendar year and calculated two measures of species richness, accounting for differences in sampling effort among years. The first measure was asymptotic species richness, *S~asymp~*, using the Chao1 estimator [@chao1984][@chao1987]. This metric is highly correlated with measured richness, *S* [@mcgill2011], but attempts to extrapolate the total species richness based on incomplete sampling. Second, we attempted to account for differences in effort among years by estimating the effort-controlled rarefied species richness, *S~n~*, which provides the expected number of species given a defined number of *n* individuals sampled [@gotelli2001]. *S~n~* estimates for yearly aggregated data, *n* was calculated by the minimum samples within a year multiplied by `r n_year`. We calculated *S~asymp~* and *S~n~* using the mobr package [@mcglinn2019][@mcglinn2021a].

@@

@@

trawl_mob_in = mobr::make_mob_in(comm = TB_trawl_taxasite %>% dplyr::select(-julian_date:-biweekly),
                           plot_attr = TB_trawl_taxasite %>% dplyr::select(julian_date:biweekly),
                           coord_names = "biweekly",
                           latlong = FALSE)

trawl_mob_stats = mobr::get_mob_stats(trawl_mob_in, group_var = 'year', index = c("N", "S", "S_asymp", "S_n", "S_PIE", "PIE", "pct_rare"), effort_samples = n_year, extrapolate = TRUE, boot_groups = TRUE, n_perm = nperm)

group_stats = trawl_mob_stats$groups_stats %>% data.frame

S_n = max(group_stats$effort, na.rm = TRUE)

#

@@

Over the sampling period from `r min(TB_trawl_sampling$year, na.rm = TRUE)` to `r max(TB_trawl_sampling$year, na.rm = TRUE)`, `r length(unique(TB_trawl_data$Date))` trawls-days were performed yielding `r format(signif(sum(TB_trawl_data$Abundance, na.rm = TRUE),3), scientific = TRUE)` individuals. Among years, total effort generally increased, both in number of trawl-days and individuals counted, from a minimum effort of `r min(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_min(trawl_days) %>% dplyr::select(year) %>% unlist`) to `r max(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_max(trawl_days) %>% dplyr::select(year) %>% unlist`) days and from `r min(TB_trawl_sampling$individuals, na.rm = TRUE)` (`r TB_trawl_sampling %>% slice_min(individuals) %>% dplyr::select(year) %>% unlist`) to `r format(signif(max(TB_trawl_sampling$individuals, na.rm = TRUE),3), scientific = TRUE)` (`r TB_trawl_sampling %>% slice_max(individuals) %>% dplyr::select(year) %>% unlist`) individuals. In total, `r length(unique(TB_trawl_data$Common_name))` unique species were identified and new species were consistently observed throughout the sampling period (Figure 1). Within a calendar year, we estimated the asymptotic species richness (*S~asymp~*) which was relatively stable for the first five years (2007 -- 2012), reaching a minimum value of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(group)`. Estimated *S~asymp~* increased in 2013 to a series maximum of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% data.frame %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(group)`. However, given the large differences in effort among years, variability in *S~asymp~* may be partly attributable to sampling effects among years. Therefore, we also calculated rarefied richness at `r S_n` individuals, *S~`r S_n`~*,  for each year. Here, effort-controlled species richness, *S~`r S_n`~*, diverged qualitatively from *S~asymp~* and was generally constant with an apparent slight increase in more recent sampling years. Overall, *S~`r S_n`~* ranged from `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(upper) %>% round(1)`) to `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(upper) %>% round(1)`) (Figure 1).

@@

TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
  dplyr::filter(!duplicated(Common_name, nmax = (total_richness) )) %>%
  dplyr::mutate(Date = as.Date(Date)) %>%
  arrange(Date) %>%
  dplyr::mutate(richness = 1:n())-> TB_time_accum

@@

### create timeseries figure
TB_time_df <- TB_time_accum %>%
  dplyr::select(Date, year, richness) %>%
  mutate(data_type = 'continuous',
         r_lower = NA_real_,
         r_upper = NA_real_) %>%
  bind_rows(TB_annual_accum %>% 
              mutate(data_type = "ann_raw",
                     r_lower = NA_real_,
                     r_upper = NA_real_)) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(effort > 100) %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "rarefied")) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(index == 'S_asymp') %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "ann_asymp")) %>%
  dplyr::mutate(data_type = factor(data_type, levels = c("continuous", "rarefied", "ann_raw", "ann_asymp"))) %>%
  dplyr::filter(data_type != "ann_raw")

@@

rich_time <-
  ggplot(TB_time_df, aes(x = Date, y = richness, group = data_type)) +
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "ann_asymp"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
              fill = viridis(n = 3, alpha = c(1,1,0.1), option = 'viridis')[3],
              inherit.aes = FALSE)+
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "rarefied"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[2],
              fill = viridis(n = 3, alpha = c(1,0.2,1), option = 'viridis')[2],
              inherit.aes = FALSE)+
  geom_segment(data = TB_time_df %>% 
                 dplyr::filter(data_type %ni% 'continuous') %>%
                 pivot_wider(Date, names_from = data_type, values_from = richness),
               aes(x = Date, xend = Date, y = ann_asymp, yend = rarefied ), 
               color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
               inherit.aes = F) +
  geom_point(data = TB_time_df, aes(colour = data_type),size = 1.5) +
  geom_line(data = TB_time_df %>% dplyr::filter(!grepl('ann_asymp', data_type)), aes(colour = data_type),size = 1.2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name = "Species Richness")+
  scale_colour_manual(labels = c("continuous", bquote(italic("S"[.(S_n[1])])),expression(italic("S"["asymp"]))), values = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')) +
  # scale_alpha_manual(values = c(1,1,0.3))+
  theme(legend.position = c(0, 1), legend.justification = c(0, 1), legend.background = element_rect(fill = "transparent", colour = NA),
        legend.title = element_blank(), legend.text.align = 0,
        # axis.title.x = element_blank(), axis.text.x = element_blank(),
        legend.spacing.y = unit(-0.1, 'cm'))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")

@@

```{r turnover plot ,echo = FALSE, warning=FALSE, message=FALSE, fig.cap = "Figure 2. Total species "}
trawl_yr_turnover %>%
  ggplot()+
  geom_line(aes(x = Date, y = value, group = variable, color = variable), size = 1.1)+
  scale_color_manual(values = viridis(n = 3, option = 'viridis'))+
  scale_y_continuous(name = 'Species turnover',limits = c(0,0.7), expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(legend.position = c(0,0),
        legend.justification = c(0,0),
        legend.background = element_rect(fill = 'transparent', color = NA),
        legend.title = element_blank(),
        axis.title.x = element_blank())
```

@@

qD_df = q2_boots %>% 
  # summarise q2 by year
  group_by(year) %>%
  dplyr::summarise(across(q2, list(mean = ~mean(.x, na.rm = TRUE),
                                   median = ~quantile(.x, 0.5, na.rm = TRUE),
                                   quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                   quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                   quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                   quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
  dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                q_order = "q2") %>%
  bind_rows(
    q1_boots %>%
      # summarise q1 by year
      group_by(year) %>%
      dplyr::summarise(across(q1, list(mean = ~mean(.x, na.rm = TRUE),
                                       median = ~quantile(.x, 0.5, na.rm = TRUE),
                                       quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                       quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                       quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                       quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
      dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                    q_order = "q1")
   )

annual_evenness_boots = boot_gini_groups(TB_trawl_taxasite %>% dplyr::select(-julian_date,-Date,-month,-biweekly), 
                                group_var = 'year',
                                n_perm = nperm)

annual_evenness_df = annual_evenness_boots %>%
  dplyr::select(-boot_id) %>%
  pivot_longer(-year, names_to = 'measure', values_to = 'value') %>%
  group_by(year,measure) %>%
  dplyr::summarise(across(value, list(mean = ~mean(.x, na.rm = TRUE),
                                      median = ~quantile(.x, 0.5, na.rm = TRUE),
                                      quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                      quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                      quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                      quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
  dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d")) %>%
  ungroup

@@

```{r evenness, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= "Figure 4." }

annual_evenness_df %>%
  dplyr::filter(measure == 'gini_raw') %>%
  ggplot()+
  geom_ribbon(aes(x = Date, ymin = quant2.5, ymax = quant97.5), alpha = 0.3) +
  geom_line(aes(x = Date, y = median), size = 1.1) +
  geom_point(aes(x = Date, y = median), shape = 22, size = 3)+
  scale_y_continuous(name = "Normalized Gini coefficient", limits = c(0,NA), expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(axis.title.x = element_blank())-> evenness_plot

group_stats %>%
  dplyr::filter(index == "pct_rare") %>%
  dplyr::mutate(group = as.Date(group, format = "%Y")) %>%
  ggplot()+
  geom_point(aes(x =group, y = median), size = 3, shape = 22)+
  geom_path(aes(x = group, y = median), size = 1.1)+
  geom_ribbon(aes(x = group, ymin = lower, ymax = upper), alpha = 0.3)+
  scale_y_continuous(name = "Percent rare species (%)", limits = c(70,100), expand = c(0.01,0.01))+
  scale_color_manual(values = viridis(n =1, option = 'viridis' ))+
  scale_fill_manual(values = viridis(n = 1, option = 'viridis'))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(axis.title.x = element_blank())-> rarity_plot

>>>>

source(here::here("datascript.R"))
source(here::here("analyses/trawl/R/Lorenz.R"))
source(here::here("analyses/trawl/R/Evenness.R"))

@@

Anthropogenic activities continue to exert long-term, varied, and increasing pressures on ecosystems causing alterations to biological diversity on broad and local scales. Shifts in biodiversity are variable, nuanced and difficult to predict. Yet, understanding how communities change through time is crucial for predicting current and future changes to biological systems and the services they provide. Explaining the extent and pace of biodiversity change will require identifying the conditions that lead to divergent biological responses, e.g., whether change happens gradually or abruptly and through individual versus cohesive species responses.

@@

<!---  --->
Anthropogenic activities continue to exert long-term, varied, and increasing pressures on ecosystems [@steffen2015] causing alterations to biological diversity on local and global scales. Shifts in biodiversity at local scales are variable and potentially more nuanced than at global scales, but generally local changes manifest primarily as changes to community composition and structure rather than declines in richness [@hillebrand2018]. These shifts in communities through time are difficult to predict, in part, due to the multidimensional nature of biodiversity [@chase2018]. Yet, understanding how communities change through time is crucial for predicting current and future changes to biological systems and ecosystem services.
<!--- --->
Explaining the extent and pace of biodiversity change will require identifying the conditions that lead to divergent biological responses to environmental change, both natural and anthropogenic. In particular, it will be important to disentangle the roles of intrinsic (e.g., ecological) and extrinsic (e.g., environmental) drivers of biodiversity change. Intrinsic and extrinsic processes can interact, leading to temporal patterns of biodiversity change that vary in response to environmental change. For example, gradual environmental change over time can lead to gradual community shifts or these shifts may be undetectable or abrupt [@williams2011][@scheffer2003].

@@

Here, we quantified biodiversity changes within a shallow estuarine community over a decade to understand shifts in total species richness, the relative distribution of abundant and rare species, and potential change points in community composition through time...

@@

Trawl sampling was conducted as part of the Louisiana Universities Marine Consortium's (LUMCON) educational outreach program (https://www.lumcon.edu/education-and-outreach). Individual trawls were conducted using *trawl apparatus description...* for 10 - 20 min. When multiple trawls were conducted on a single day, species counts were summed across all trawls...

@@

Species names were checked for typos and confirmed using the *gni_parse* function from the *taxize* package [@chamberlain2020] and further standardized to consistent taxonomic resolution for entire data series (Supplemental materials). We measured multiple aspects of species richness, first to track the uncorrected accumulation of species over time we filtered species from the data set after first occurrence. We then aggregated all trawls within a calendar year and calculated two measures of species richness, accounting for differences in sampling effort among years. The first measure was asymptotic species richness, *S~asymp~*, using the Chao1 estimator [@chao1984][@chao1987]. This metric is highly correlated with measured richness, *S* [@mcgill2011], but attempts to extrapolate the total species richness based on incomplete sampling. Second, we attempted to account for differences in effort among years by estimating the effort-controlled rarefied species richness, *S~n~*, which provides the expected number of species given a defined number of *n* individuals sampled [@gotelli2001]. *S~n~* estimates for yearly aggregated data, *n* was calculated by the minimum number of trawls within a year multiplied by `r n_year`. We calculated *S~asymp~* and *S~n~* using the mobr package [@mcglinn2019][@mcglinn2021a].

@@

TB_trawl_taxasite_yr <- TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
  dplyr::select(year, Common_name, Abundance) %>%
  group_by(year, Common_name) %>%
  summarise(Abundance = sum(Abundance, na.rm = TRUE)) %>%
  na.omit %>% #gropu_by(year) %>%
  # remove trawls with less than 20 individuals 
  # dplyr::filter(sum(Abundance) >=20) %>%
  group_by(year, Common_name) %>%
  pivot_wider(names_from = Common_name, values_from = Abundance, values_fill = list(Abundance = 0)) %>%
  ungroup() %>%
  dplyr::select( year, everything())

@@

## Estimated mobr for biweekly  for all communities
# TB_trawl_biweekly = TB_trawl_data %>%
#   # exclude 2020 & 2021
#   dplyr::filter(year %ni% c("2020","2021")) %>%
#   dplyr::mutate(julian_date = julian(Date, origin = as.Date("2007-01-05")),
#                 biweekly = ceiling(julian_date/14)) %>%
#   group_by(year,biweekly, Common_name) %>%
#   dplyr::summarise(Abundance = sum(Abundance, na.rm = TRUE)) %>%
#   group_by(year,biweekly) %>%
#   pivot_wider( names_from = 'Common_name', values_from = 'Abundance', values_fill = 0) %>%
#   ungroup

# trawl_mob_in = mobr::make_mob_in(comm = TB_trawl_biweekly %>% dplyr::select(-biweekly),
#                                  plot_attr = TB_trawl_biweekly %>% dplyr::select(year,biweekly),
#                                  coord_names = "biweekly",
#                                  latlong = FALSE)

# trawl_mob_in = mobr::make_mob_in(comm = TB_trawl_taxasite %>% dplyr::select(-julian_date:-biweekly),
#                                  plot_attr = TB_trawl_taxasite %>% dplyr::select(julian_date:biweekly),
#                                  coord_names = "biweekly",
#                                  latlong = FALSE)

# trawl_mob_stats = mobr::get_mob_stats(trawl_mob_in, group_var = 'year', index = c("N", "S", "S_asymp", "S_n", "S_PIE", "PIE", "pct_rare"), effort_samples = n_year, extrapolate = TRUE, boot_groups = TRUE, n_perm = nperm)

# group_stats = trawl_mob_stats$groups_stats %>% data.frame

chao1_yr = TB_trawl_taxasite_yr[,-1] %>%
  apply(., 1, mobr::calc_chao1) %>%
  bind_cols(TB_trawl_taxasite_yr$year,.) %>%
  setNames(., c("year","S_asymp")) %>%
  dplyr::mutate(Date = as.Date(paste0(year,"-01-01"), format = "%Y-%m-%d"))

S_n = min(TB_trawl_sampling$trawl_days, na.rm = TRUE)*n_year

Sn_yr = TB_trawl_taxasite_yr[,-1] %>%
  apply(., 1, function(x) mobr::rarefaction(x, effort = S_n, method = "IBR", extrapolate = TRUE)) %>%
  bind_cols(TB_trawl_taxasite_yr$year,.) %>%
  setNames(., c("year","S_n")) %>%
  dplyr::mutate(Date = as.Date(paste0(year,"-01-01"), format = "%Y-%m-%d"))

pct_rare =TB_trawl_taxasite_yr[,-1] %>%
  apply(., 1, function(x) pct_rare(x, 0.05)) %>%
  bind_cols(TB_trawl_taxasite_yr$year,.) %>%
  setNames(., c("year","pct_rare")) %>%
  dplyr::mutate(Date = as.Date(paste0(year,"-01-01"), format = "%Y-%m-%d"))

group_stats = left_join(chao1_yr, Sn_yr) %>%
  left_join(pct_rare) %>%
  dplyr::select(Date, year, everything()) %>%
  group_by(Date, year) %>%
  pivot_longer(c(S_n,S_asymp,pct_rare),names_to = "index", values_to = "median" ) %>%
  ungroup
#

@@

Over the sampling period from `r min(TB_trawl_sampling$year, na.rm = TRUE)` to `r max(TB_trawl_sampling$year, na.rm = TRUE)`, `r length(unique(TB_trawl_data$Date))` trawls-days were performed yielding `r format(signif(sum(TB_trawl_data$Abundance, na.rm = TRUE),3), scientific = TRUE)` individuals. Among years, total effort generally increased, both in number of trawl-days and individuals counted, from a minimum effort of `r min(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_min(trawl_days) %>% dplyr::select(year) %>% unlist`) to `r max(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_max(trawl_days) %>% dplyr::select(year) %>% unlist`) days and from `r min(TB_trawl_sampling$individuals, na.rm = TRUE)` (`r TB_trawl_sampling %>% slice_min(individuals) %>% dplyr::select(year) %>% unlist`) to `r format(signif(max(TB_trawl_sampling$individuals, na.rm = TRUE),3), scientific = TRUE)` (`r TB_trawl_sampling %>% slice_max(individuals) %>% dplyr::select(year) %>% unlist`) individuals. In total, `r length(unique(TB_trawl_data$Common_name))` unique species were identified and new species were consistently observed throughout the sampling period (Figure 1). Within a calendar year, we estimated the asymptotic species richness (*S~asymp~*) which was relatively stable for the first five years (2007 -- 2012), reaching a minimum value of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` in `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(year)`. Estimated *S~asymp~* increased in 2013 to a series maximum of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` in `r group_stats %>% data.frame %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(year)`. However, given the large differences in effort among years, variability in *S~asymp~* may be partly attributable to sampling effects among years. Therefore, we also calculated rarefied richness at `r S_n` individuals, *S~`r S_n`~*,  for each year. Here, effort-controlled species richness, *S~`r S_n`~*, diverged qualitatively from *S~asymp~* and was generally constant with an apparent slight increase in more recent sampling years. Overall, *S~`r S_n`~* ranged from `r group_stats %>% dplyr::filter(index == "S_n") %>% slice_min(median) %>% dplyr::select(median) %>% round(1)` to `r group_stats %>% dplyr::filter(index == "S_n") %>% slice_max(median) %>% dplyr::select(median) %>% round(1)` (Figure 1).

@@

TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
  dplyr::filter(!duplicated(Common_name )) %>%
  dplyr::mutate(Date = as.Date(Date)) %>%
  arrange(Date) %>%
  dplyr::mutate(richness = 1:n())-> TB_time_accum

@@

### create timeseries figure
TB_time_df <- TB_time_accum %>%
  dplyr::select(Date, year, richness) %>%
  mutate(data_type = 'continuous') %>%
  bind_rows(TB_annual_accum %>% 
              mutate(data_type = "ann_raw")) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(index == "S_n") %>%
              dplyr::select(Date, year, richness = 'median') %>%
              dplyr::mutate(data_type = "rarefied")) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(index == 'S_asymp') %>%
              dplyr::select(Date, year, richness = 'median') %>%
              dplyr::mutate(data_type = "ann_asymp")) %>%
  dplyr::mutate(data_type = factor(data_type, levels = c("continuous", "rarefied", "ann_raw", "ann_asymp"))) %>%
  dplyr::filter(data_type != "ann_raw")

@@

rich_time <-
  ggplot(TB_time_df, aes(x = Date, y = richness, group = data_type)) +
  geom_segment(data = TB_time_df %>% 
                 dplyr::filter(data_type %ni% 'continuous') %>%
                 pivot_wider(Date, names_from = data_type, values_from = richness),
               aes(x = Date, xend = Date, y = ann_asymp, yend = rarefied ), 
               color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
               inherit.aes = F) +
  geom_point(data = TB_time_df, aes(colour = data_type),size = 1.5) +
  geom_line(data = TB_time_df %>% dplyr::filter(!grepl('ann_asymp', data_type)), aes(colour = data_type),size = 1.2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",limits = c(as.Date("2007-01-01"),as.Date("2019-06-01"))) +
  scale_y_continuous(name = "Species Richness")+
  scale_colour_manual(labels = c("continuous", bquote(italic("S"[.(S_n[1])])),expression(italic("S"["asymp"]))), values = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')) +
  # scale_alpha_manual(values = c(1,1,0.3))+
  theme(legend.position = c(0, 1), legend.justification = c(0, 1), legend.background = element_rect(fill = "transparent", colour = NA),
        legend.title = element_blank(), legend.text.align = 0,
        # axis.title.x = element_blank(), axis.text.x = element_blank(),
        legend.spacing.y = unit(-0.1, 'cm'))

@@

```{r turnover plot ,echo = FALSE, warning=FALSE, message=FALSE, fig.cap = "Figure 2. Total species turnover and partitioning"}
trawl_yr_turnover %>%
  ggplot()+
  geom_line(aes(x = Date, y = value, group = variable, color = variable), size = 1.1)+
  scale_color_manual(values = viridis(n = 3, option = 'viridis'))+
  scale_y_continuous(name = 'Species turnover',limits = c(0,0.7), expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(legend.position = c(0,0),
        legend.justification = c(0,0),
        legend.background = element_rect(fill = 'transparent', color = NA),
        legend.title = element_blank(),
        axis.title.x = element_blank())
```

@@

qD_df = q2_boots %>% 
  # summarise q2 by year
  group_by(year) %>%
  dplyr::summarise(across(q2, list(mean = ~mean(.x, na.rm = TRUE),
                                   median = ~quantile(.x, 0.5, na.rm = TRUE),
                                   quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                   quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                   quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                   quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
  dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                q_order = "q2") %>%
  bind_rows(
    q1_boots %>%
      # summarise q1 by year
      group_by(year) %>%
      dplyr::summarise(across(q1, list(mean = ~mean(.x, na.rm = TRUE),
                                       median = ~quantile(.x, 0.5, na.rm = TRUE),
                                       quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
                                       quant25= ~quantile(.x, 0.25, na.rm = TRUE),
                                       quant75= ~quantile(.x, 0.75, na.rm = TRUE),
                                       quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
      dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                    q_order = "q1")
   ) 
# 
# annual_evenness_boots = boot_gini_groups(TB_trawl_taxasite %>% dplyr::select(-julian_date,-Date,-month,-biweekly), 
#                                 group_var = 'year',
#                                 n_perm = nperm)
# 
# annual_evenness_df = annual_evenness_boots %>%
#   dplyr::select(-boot_id) %>%
#   pivot_longer(-year, names_to = 'measure', values_to = 'value') %>%
#   group_by(year,measure) %>%
#   dplyr::summarise(across(value, list(mean = ~mean(.x, na.rm = TRUE),
#                                       median = ~quantile(.x, 0.5, na.rm = TRUE),
#                                       quant2.5= ~quantile(.x, 0.025, na.rm = TRUE),
#                                       quant25= ~quantile(.x, 0.25, na.rm = TRUE),
#                                       quant75= ~quantile(.x, 0.75, na.rm = TRUE),
#                                       quant97.5= ~quantile(.x, 0.975, na.rm = TRUE)), .names = "{.fn}")) %>%
#   dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d")) %>%
#   ungroup

annual_evenness_df = TB_trawl_taxasite_yr %>%
  dplyr::select(-year) %>%
  apply(., 1, gini_even) %>%
      t() %>%
      data.frame() %>%
      bind_cols(TB_trawl_taxasite_yr$year, .) %>%
      setNames(., c("year", "gini_raw", "gini_norm")) %>%
  pivot_longer(-year, names_to = "measure", values_to = "median") %>%
  dplyr::mutate(Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d")) %>%
  ungroup

@@

```{r evenness, echo = FALSE, warning=FALSE, message=FALSE, fig.cap= "Figure 4. Species rarity and evenness in relative abundances through the data series." }

annual_evenness_df %>%
  dplyr::filter(measure == 'gini_norm') %>%
  ggplot()+
  # geom_ribbon(aes(x = Date, ymin = quant2.5, ymax = quant97.5), alpha = 0.3) +
  geom_line(aes(x = Date, y = median), size = 1.1) +
  geom_point(aes(x = Date, y = median), shape = 22, size = 3)+
  scale_y_continuous(name = "Normalized Gini coefficient", limits = c(0,0.2), expand = c(0.01,0.01))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(axis.title.x = element_blank())-> evenness_plot

group_stats %>%
  dplyr::filter(index == "pct_rare") %>%
  dplyr::mutate(year = as.Date(year, format = "%Y")) %>%
  ggplot()+
  geom_point(aes(x =Date, y = median), size = 3, shape = 22)+
  geom_path(aes(x = Date, y = median), size = 1.1)+
  scale_y_continuous(name = "Percent rare species (%)", limits = c(75 ,100), expand = c(0.01,0.01))+
  scale_color_manual(values = viridis(n =1, option = 'viridis' ))+
  scale_fill_manual(values = viridis(n = 1, option = 'viridis'))+
  scale_x_date(date_breaks = 'year', limits = c(as.Date("2007-01-01"),as.Date("2019-06-01")),
               date_labels = "%Y")+
  theme(axis.title.x = element_blank())-> rarity_plot
